### 2020-10-13 priority-gv.tex
Minor update.
Changed float position.
Finished PGV section.

\section{Priority GV}
\usingnamespace{pgv}

@@

\paragraph{Types}
Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
%
Session types are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{S}
  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
\end{array}
\]
The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$. The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel. Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.

@@

\paragraph{Environments}
Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types. Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
\[
\begin{array}{lcl}
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]

\paragraph{Duality}
Duality plays a crucial role in session types. The two endpoints of a channel are assigned dual types, which ensures that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}. Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$. Duality is an involutive function which preserves priorities.
\[
\begin{array}{lcl}
  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
\end{array}
\]

\paragraph{Priorities}
The function $\pr$ returns the smallest priority of a session type. The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
\[
\begin{array}{lcl}
  \pr(\ty{\tysend[\cs{o}]{T}{S}})            & = & \cs{o} \\
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}})            & = & \cs{o} \\
\end{array}
\qquad
\begin{array}{lcl}
  \pr(\ty{\tyends[\cs{o}]})                  & = & \cs{o} \\
  \pr(\ty{\tyendr[\cs{o}]})                  & = & \cs{o} \\
\end{array}
\]
We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority:
\[
\begin{array}{lcl}
  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \minpr(\ty{\tyunit})                       & = & \ptop \\
  \minpr(\ty{\tyvoid})                       & = & \ptop \\
  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
  \minpr(\ty{\emptyenv})                     & = & \ptop
\end{array}
\]

\paragraph{Terms}
Terms ($\tm{L}$, $\tm{M}$, $\tm{N}$) are defined by the following grammar:
\[
\begin{array}{lcl}
  \tm{L}, \tm{M}, \tm{N}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{M\;N} \\
  & \sep      & \tm{\unit}
    \sep        \tm{\letunit{M}{N}} \\
  & \sep      & \tm{\pair{M}{N}}
    \sep        \tm{\letpair{x}{y}{M}{N}} \\
  & \sep      & \tm{\inl{M}}
    \sep        \tm{\inr{M}}
    \sep        \tm{\casesum{L}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{M}} \\
  \tm{K}
  & \coloneqq & \tm{\link}
    \sep        \tm{\new}
    \sep        \tm{\spawn}
    \sep        \tm{\send}
    \sep        \tm{\recv}
    \sep        \tm{\close}
    \sep        \tm{\wait}
\end{array}
\]
Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ range over variable names. Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$, which range over \emph{free} endpoints. The term language is the standard linear $\lambda$-calculus with products, sums, and their units, extended with constants $\tm{K}$ for the communication primitives.

@@

\paragraph{Internal and External Choice}
Typically, session-typed languages feature constructs for internal and external choice. In GV, these can be defined in terms of the core language, by sending or receiving a value of a sum type~\cite{lindleymorris15}. We use the following syntactic sugar for internal ($\ty{\tyselect[\cs{o}]{S}{S'}}$) and external ($\ty{\tyoffer[\cs{o}]{S}{S'}}$) choice and their units:
\[
\begin{array}{lcl}
  \ty{\tyselect[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
  \ty{\tyoffer[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\tyselectemp[\cs{o}]}
  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
  \ty{\tyofferemp[\cs{o}]}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
\end{array}
\]
As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types. For simplicity, we present only the binary and nullary forms.

@@

\paragraph{Operational Semantics}
Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
\[
\begin{array}[t]{lcl}
    \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  & \coloneqq & \tm{\phi\;M}
    \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
    \sep        \tm{\res{x}{x'}{\conf{C}}}
  \\
  \tm{\phi}
  & \coloneqq & \tm{\main}
    \sep        \tm{\child}
\end{array}
\]
Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$. To preserve the functional nature of PGV, where programs return a single value, we use flags ($\tm{\phi}$) to differentiate between the main thread, marked $\tm{\main}$, and child threads created by $\tm{\spawn}$, marked $\tm{\child}$. Only the main thread returns a value. We determine the flag of a configuration by combining the flags of all threads in that configuration:
\[
\begin{array}{ll}
  \tm{\main}  + \tm{\child} = \tm{\main}
  \\
  \tm{\child} + \tm{\main}  = \tm{\main}
\end{array}
\quad
\begin{array}{ll}
  \tm{\child} + \tm{\child} = \tm{\child}
  \\
  \tm{\main}  + \tm{\main} \; \text{undefined}
\end{array}
\]
Values ($\tm{V}$, $\tm{W}$), ready terms ($\tm{R}$) evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
\[
\begin{array}[t]{lcl}
  \tm{V}, \tm{W}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{\unit}
    \sep        \tm{\pair{M}{N}}
    \sep        \tm{\inl{M}}
    \sep        \tm{\inr{M}}
  \\
  \tm{E}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{E\;M}
    \sep        \tm{V\;E} \\
  & \sep      & \tm{\letunit{E}{N}} \\
  & \sep      & \tm{\pair{E}{M}}
    \sep        \tm{\pair{V}{E}}
    \sep        \tm{\letpair{x}{y}{E}{M}} \\
  & \sep      & \tm{\inl{E}}
    \sep        \tm{\inr{E}}
    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{E}}
  \\
  \tm{\conf{F}}
  & \coloneqq & \tm{\phi\;E}
  \\
  \tm{\conf{G}}
  & \coloneqq & \tm{\hole}
    \sep        \tm{\ppar{\conf{G}}{\conf{C}}}
    \sep        \tm{\res{x}{y}{\conf{G}}}
\end{array}
\]
We factor the reduction relation of PGV into a deterministic reduction on terms ($\tred$) and a non-deterministic reduction on configurations ($\cred$), see \cref{fig:pgv-operational-semantics}. We write $\tred^+$ and $\cred^+$ for the transitive closures, and $\tred^\star$ and $\cred^\star$ for the reflexive-transitive closures.

@@

\paragraph{Typing}
\Cref{fig:pgv-typing} gives the typing rules for PGV.

@@

\paragraph{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}. Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.

@@

\subsection{Metatheory}

\begin{restatablelemma}{lempgvvaluedone}
  \label{lem:pgv-value-done}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
  Details in \cref{prf:lem-pgv-value-done}.
\end{proof}

\begin{restatablelemma}{lempgvsubstitution}[Substitution]
  \label{lem:pgv-substitution}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$. Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved.
  Details in \cref{prf:lem-pgv-substitution}.
\end{proof}

\begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
  \label{lem:pgv-subject-reduction-terms}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
  then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
  Details in \cref{prf:lem-pgv-subject-reduction-terms}.
\end{proof}

\begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
  \label{lem:pgv-subject-congruence}
  \hfill\\%newline before theorem statement
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{\conf{C}}\equiv\tm{\conf{C'}}$.
  Details in \cref{prf:lem-pgv-subject-congruence}.
\end{proof}

\begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
  \label{thm:pgv-subject-reduction-confs}
  \hfill\\%newline before theorem statement
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatabletheorem}
\begin{proof}
  By induction on the derivation of $\tm{\conf{C}}\cred\tm{\conf{C'}}$.
  Details in \cref{prf:thm-pgv-subject-reduction-confs}.
\end{proof}

\begin{definition}[Actions]
  \label{def:pgv-actions}
  A~term acts on an endpoint $\tm{x}$ if it is of the form $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$. A~term is an action if it acts on some endpoint $\tm{x}$.
\end{definition}

\begin{definition}[Ready Terms]
  \label{def:pgv-ready-actions}
  A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$. In the latter case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
\end{definition}

\begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
  \label{lem:pgv-open-progress-terms}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, then:
  \begin{itemize}
  \item $\tm{M}$ is a value;
  \item $\tm{M}\tred\tm{N}$ for some term $\tm{N}$; or
  \item $\tm{M}$ is ready.
  \end{itemize}
\end{restatablelemma}

\begin{definition}[Canonical Forms]
  \label{def:pgv-canonical-forms}
  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form, where no term $\tm{M_i}$ is a value:
  \[
    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
  \]
\end{definition}

\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
  \label{lem:pgv-canonical-forms}
  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
\end{restatablelemma}
\begin{proof}
  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
\end{proof}

>>>>

\section{Priority GV}\label{sec:pgv}
\usingnamespace{pgv}

@@

\subsubsection*{Types}
Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
%
Session types are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{S}
  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
\end{array}
\]
The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$. The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel. Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.

@@

\subsubsection*{Environments}
Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types. Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
\[
\begin{array}{lcl}
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]

\subsubsection*{Duality}
Duality plays a crucial role in session types. The two endpoints of a channel are assigned dual types, which ensures that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}. Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$. Duality is an involutive function which preserves priorities.
\[
\begin{array}{lcl}
  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
\end{array}
\]

\subsubsection*{Priorities}
The function $\pr$ returns the smallest priority of a session type. The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
\[
\begin{array}{lcl}
  \pr(\ty{\tysend[\cs{o}]{T}{S}})            & = & \cs{o} \\
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}})            & = & \cs{o} \\
\end{array}
\qquad
\begin{array}{lcl}
  \pr(\ty{\tyends[\cs{o}]})                  & = & \cs{o} \\
  \pr(\ty{\tyendr[\cs{o}]})                  & = & \cs{o} \\
\end{array}
\]
We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority:
\[
\begin{array}{lcl}
  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \minpr(\ty{\tyunit})                       & = & \ptop \\
  \minpr(\ty{\tyvoid})                       & = & \ptop \\
  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
  \minpr(\ty{\emptyenv})                     & = & \ptop
\end{array}
\]

\subsubsection*{Terms}
Terms ($\tm{L}$, $\tm{M}$, $\tm{N}$) are defined by the following grammar:
\[
\begin{array}{lcl}
  \tm{L}, \tm{M}, \tm{N}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{M\;N} \\
  & \sep      & \tm{\unit}
    \sep        \tm{\letunit{M}{N}} \\
  & \sep      & \tm{\pair{M}{N}}
    \sep        \tm{\letpair{x}{y}{M}{N}} \\
  & \sep      & \tm{\inl{M}}
    \sep        \tm{\inr{M}}
    \sep        \tm{\casesum{L}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{M}} \\
  \tm{K}
  & \coloneqq & \tm{\link}
    \sep        \tm{\new}
    \sep        \tm{\spawn}
    \sep        \tm{\send}
    \sep        \tm{\recv}
    \sep        \tm{\close}
    \sep        \tm{\wait}
\end{array}
\]
Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ range over variable names. Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$, which range over \emph{free} endpoints. The term language is the standard linear $\lambda$-calculus with products, sums, and their units, extended with constants $\tm{K}$ for the communication primitives.

@@

\subsubsection*{Internal and External Choice}
Typically, session-typed languages feature constructs for internal and external choice. In GV, these can be defined in terms of the core language, by sending or receiving a value of a sum type~\cite{lindleymorris15}. We use the following syntactic sugar for internal ($\ty{\tyselect[\cs{o}]{S}{S'}}$) and external ($\ty{\tyoffer[\cs{o}]{S}{S'}}$) choice and their units:
\[
\begin{array}{lcl}
  \ty{\tyselect[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
  \ty{\tyoffer[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\tyselectemp[\cs{o}]}
  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
  \ty{\tyofferemp[\cs{o}]}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
\end{array}
\]
As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types. For simplicity, we present only the binary and nullary forms.

@@

\subsubsection*{Operational Semantics}
Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
\[
\begin{array}[t]{lcl}
    \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  & \coloneqq & \tm{\phi\;M}
    \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
    \sep        \tm{\res{x}{x'}{\conf{C}}}
  \\
  \tm{\phi}
  & \coloneqq & \tm{\main}
    \sep        \tm{\child}
\end{array}
\]
Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$. To preserve the functional nature of PGV, where programs return a single value, we use flags ($\tm{\phi}$) to differentiate between the main thread, marked $\tm{\main}$, and child threads created by $\tm{\spawn}$, marked $\tm{\child}$. Only the main thread returns a value. We determine the flag of a configuration by combining the flags of all threads in that configuration:
\[
\begin{array}{ll}
  \tm{\main}  + \tm{\child} = \tm{\main}
  \\
  \tm{\child} + \tm{\main}  = \tm{\main}
\end{array}
\quad
\begin{array}{ll}
  \tm{\child} + \tm{\child} = \tm{\child}
  \\
  \tm{\main}  + \tm{\main} \; \text{undefined}
\end{array}
\]
Values ($\tm{V}$, $\tm{W}$), ready terms ($\tm{R}$) evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
\[
\begin{array}[t]{lcl}
  \tm{V}, \tm{W}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{\unit}
    \sep        \tm{\pair{M}{N}}
    \sep        \tm{\inl{M}}
    \sep        \tm{\inr{M}}
  \\
  \tm{E}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{E\;M}
    \sep        \tm{V\;E} \\
  & \sep      & \tm{\letunit{E}{N}} \\
  & \sep      & \tm{\pair{E}{M}}
    \sep        \tm{\pair{V}{E}}
    \sep        \tm{\letpair{x}{y}{E}{M}} \\
  & \sep      & \tm{\inl{E}}
    \sep        \tm{\inr{E}}
    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{E}}
  \\
  \tm{\conf{F}}
  & \coloneqq & \tm{\phi\;E}
  \\
  \tm{\conf{G}}
  & \coloneqq & \tm{\hole}
    \sep        \tm{\ppar{\conf{G}}{\conf{C}}}
    \sep        \tm{\res{x}{y}{\conf{G}}}
\end{array}
\]
We factor the reduction relation of PGV into a deterministic reduction on terms ($\tred$) and a non-deterministic reduction on configurations ($\cred$), see \cref{fig:pgv-operational-semantics}. We write $\tred^+$ and $\cred^+$ for the transitive closures, and $\tred^\star$ and $\cred^\star$ for the reflexive-transitive closures.

@@

\subsubsection*{Typing}
\Cref{fig:pgv-typing} gives the typing rules for PGV.

@@

\subsubsection*{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}. Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.

@@

\subsubsection*{Subject reduction}
Unlike with previous versions of GV, The structural congruence, term reduction, and configuration reduction are all type preserving, though it is not entirely trivial to show.

Specifically, we need to show that substitution preserves priority constraints. For this, it helps to prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication actions, and that any priorities present in the type of the value come from the typing environment.
\begin{restatablelemma}{lempgvvaluedone}
  \label{lem:pgv-value-done}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$. We demonstrate the case for \LabTirName{T-Inl}, as it illustrates the need for the side conditions on \LabTirName{T-Inl} and \LabTirName{T-Inr}:
  \begin{case*}[\LabTirName{T-Inl}]
    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{T})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
    \begin{mathpar}
      \small
      \inferrule*{
        \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
        \\
        \pr(\ty{T})=\pr(\ty{U})
      }{\tseq[\cs{p}]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
    \end{mathpar}
  \end{case*}
  The full proof can be found in~\cref{prf:lem-pgv-value-done}.
\end{proof}
\begin{restatablelemma}{lempgvsubstitution}[Substitution]
  \label{lem:pgv-substitution}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$. Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved. We demonstrate the case for \LabTirName{T-Lam}:
  \begin{case*}[\LabTirName{T-Lam}]
    By \cref{lem:pgv-value-done}, $\pr(\ty{\Theta})=\pr(\ty{U'})$, hence $\pr(\ty{\Gamma},\ty{\Theta})=\pr(\ty{\Gamma},\ty{U'})$.
    \begin{mathpar}
      \small
      \inferrule*{
        \tseq[\cs{q}]{\ty{\Gamma},\tmty{x}{U'},\tmty{y}{T}}{M}{U}
      }{\tseq[\cs{\pbot}]{\ty{\Gamma},\tmty{x}{U'}}
        {\lambda y.M}
        {\tylolli[\cs{\pr(\ty{\Gamma},\ty{U'})},\cs{q}]{T}{U}}}
      \substarrow{V}{x}
      \inferrule*{
        \tseq[\cs{q}]{\ty{\Gamma},\ty{\Theta},\tmty{y}{T}}{\subst{M}{V}{x}}{U}
      }{\tseq[\cs{\pbot}]{\ty{\Gamma},\ty{\Theta}}
        {\lambda y.\subst{M}{V}{x}}
        {\tylolli[\cs{\pr(\ty{\Gamma},\ty{\Theta})},\cs{q}]{T}{U}}}
    \end{mathpar}
  \end{case*}
  The complete proof can be found in~\cref{prf:lem-pgv-substitution}.
\end{proof}
The proofs for subject congruence and subject reduction are standard.
\begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
  \label{lem:pgv-subject-reduction-terms}
  \hfill\\%newline before theorem statement
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
  then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
  Details in \cref{prf:lem-pgv-subject-reduction-terms}.
\end{proof}
\begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
  \label{lem:pgv-subject-congruence}
  \hfill\\%newline before theorem statement
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatablelemma}
\begin{proof}
  By induction on the derivation of $\tm{\conf{C}}\equiv\tm{\conf{C'}}$.
  Details in \cref{prf:lem-pgv-subject-congruence}.
\end{proof}
\begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
  \label{thm:pgv-subject-reduction-confs}
  \hfill\\%newline before theorem statement
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatabletheorem}
\begin{proof}
  By induction on the derivation of $\tm{\conf{C}}\cred\tm{\conf{C'}}$.
  Details in \cref{prf:thm-pgv-subject-reduction-confs}.
\end{proof}
